A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe

### 2018-12-02 02_Materials_and_Methods.Rmd
more work on materials&methods

To test for correlation between the distances based on the variables *burial type* and *burial construction*, the timeframe of interest was devided into sections of 200 years. For each section of each the two variables, a matrix of the mean SED over time was calculated. The two matrizes of each section can be tested for correlation with the Mantel test, which **TODO** [@mantelDetectionDiseaseClustering1967]. Since both distances compared here, *burial type* and *burial construction* distance, are interval (**?**) scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test. The mean correlation for each of the $\frac{8*8}{2}-\frac{8}{2}=28$ region-region relations can be plotted in the *burial type*SED-*burial construction*SED space (Figure **TODO**).

Defining a measure of spatial distance for the above defined main units of observation, the macroregions, is only possible in a highly simplified way. A contionus distance measure can't be employed, but it's possible to distinguish ordinal distance classes. For the eight regions four different categories of spatial distance were defined.

The Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] because of the difference in the level of measurement of cultural and spatial distance.

To get a better understanding how the correlation coefficient should develop if there were strong correlation between spatial and cultural distance as measured by *burial type* and *burial construction*, a simple simulation was implemented based on the model proposed by @neimanStylisticVariationEvolutionary1995. This simulation assumes a population separated into $g$ groups with a group population of $N_g$ each. At the beginning of the simulation there are $k$ ideas randomly, but equally distributed in this population. With every timestep -- $t$ defines the total number of timesteps -- all individuals learn from an other randomly chosen individual. The probability that this other individual is from an other group equals $m_i$, for the opposite therefore $1-m_i$. The partner group for each individual learning instance is selected randomly following a group interaction probability defined in the matrix $I$.

>>>>

To test for correlation between the distances based on the variables *burial type* and *burial construction*, the timeframe of interest was devided into sections of 200 years. For each section of each the two variables, a matrix of the mean SED over time was calculated. The two matrizes of each section can be tested for correlation with the Mantel test, which **TODO** [@mantelDetectionDiseaseClustering1967]. Since both distances compared here, *burial type* and *burial construction* distance, are cardinal scale variables, the Pearson correlation coefficient [@pearsonNotesRegressionInheritance1895] can be applied as a base measure for the Mantel test. The mean correlation for each of the $\frac{8*8}{2}-\frac{8}{2}=28$ region-region relations can be plotted in the *burial type*SED-*burial construction*SED space (Figure **TODO**).

Defining a measure of spatial distance for the above established main units of observation, the macroregions, is only possible in a highly simplified way. A contionus distance measure can't be employed, but it's possible to distinguish ordinal distance classes. For the eight regions four different categories of spatial distance were defined. The test for correlation between spatial and cultural distance works similar to the one between the two cultural distances above, except that the Mantel test has to be applied with Spearman's rank correlation coefficient [@spearmanProofMeasurementAssociation1904] here, because of the difference in the data level of measurements (Figure **TODO**).

To get a better understanding how the correlation coefficient should develop if there were strong correlation between spatial and cultural distance as measured by *burial type* and *burial construction*, a simple simulation was implemented based on the model proposed by @neimanStylisticVariationEvolutionary1995. This simulation assumes a population separated into $g$ groups with a group population of $N_g$ each. At the beginning of the simulation there are $k$ ideas randomly, but equally distributed in this population. With every timestep -- $t$ defines the total number of timesteps -- all individuals learn from an other randomly chosen individual. The probability that this other individual is from an other group equals $m_i$, for the opposite therefore $1-m_i$. The partner group for each individual learning instance is selected randomly following a group interaction probability defined in the matrix $I$. The distribution of the $k$ variants can be observed for every timestep and documented in a timeseries like the one for real world data (Figure **TODO**).

For the actual simulation runs, the parameters were chosen as follows: **TODO**
